\ieeetitle

### 2016-12-27 report/figures/architecture.tex
WIP report

	\begin{tikzpicture}[transform shape,scale=0.6]
\node[draw, text width=2.1cm, text centered] (bench) at (2.1, 2.6) {Benchmarks runner};
\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python script}};
\draw[dotted,red,thick] (0.8,2) rectangle (3.4,3.4);
%
\draw[loosely dashed,black] (3.7,4.3) rectangle (14.8,-0.8);
\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay:} \texttt{epto containers communicate between each other using UDP}};
%
\node[draw] (tck) at (12.4, 2.5) {Tracker};
\draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto-tracker:}};
\draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);
\draw[dotted,red,thick] (11.6,2) rectangle (14.4,3.25);
%
\node[draw] (app1) at (7.1, 2.5) {EpTOTester 1};
\draw (5.7,3.65) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (5.8,3.25) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (5.7,2) rectangle (8.5,3.65);
\draw[dotted,red,thick] (5.8,2.1) rectangle (8.4,3.25);
%
\node[draw] (app2) at (5.4, 0) {EpTOTester 2};
\draw (4,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (4,-0.5) rectangle (6.8,1.15);
\draw[dotted,red,thick] (4.1,-0.4) rectangle (6.7,.75);
%
\node[draw] (appn) at (9.4, 0) {EpTOTester N};
\draw (8,1.15) node[below right,blue,scale=0.8] {\textit{Docker:} \texttt{epto}};
\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[loosely dashed,blue] (8,-0.5) rectangle (10.8,1.15);
\draw[dotted,red,thick] (8.1,-0.4) rectangle (10.7,.75);
%
\draw[<->,thick,>=stealth] (app1.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (app2.east) to[out=50,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (appn.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
%
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app1.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (app2.west);
\draw[<->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Start/Stop} (appn.west);
\end{tikzpicture}

>>>>

\begin{tikzpicture}[transform shape,scale=0.6]
\node[draw, text width=2.1cm, text centered] (bench) at (2.1, 2.6) {Benchmarks runner};
\draw (0.8,3.4) node[below right,red,scale=0.8] {\textit{Python script}};
\draw[dotted,red,thick] (0.8,2) rectangle (3.4,3.4);
%
\draw[loosely dashed,black] (3.7,4.3) rectangle (14.8,-0.8);
\draw (3.7,4.3) node[below right,black,scale=0.8] {\textit{Swarm overlay}};
%
\node[draw] (tck) at (12.4, 2.5) {Tracker};
\draw (11.6,3.25) node[below right,red,scale=0.8] {\textit{Python script}};
\draw (11.5,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto-tracker:}};
\draw[loosely dashed,blue] (11.5,1.9) rectangle (14.5,3.65);
\draw[dotted,red,thick] (11.6,2) rectangle (14.4,3.25);
%
\draw[loosely dashed,blue] (4,-0.5) rectangle (10.8,3.65);
\draw (4,3.65) node[below right,blue,scale=0.8] {\textit{Service:} \texttt{epto}};
%
\node[draw] (app1) at (7.1, 2.5) {EpTOTester 1};
\draw (5.8,3.25) node[below right,red,scale=0.8] {\textit{Java program}};
\draw[dotted,red,thick] (5.8,2.1) rectangle (8.4,3.25);
%
\node[draw] (app2) at (5.4, 0) {EpTOTester 2};
\draw[dotted,red,thick] (4.1,-0.4) rectangle (6.7,.75);
%
\node[draw] (appn) at (9.4, 0) {EpTOTester N};
\draw[dotted,red,thick] (8.1,-0.4) rectangle (10.7,.75);
%
\draw[<->,thick,>=stealth] (app1.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (app2.east) to[out=55,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
\draw[<->,thick,>=stealth] (appn.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{HTTP} (tck.west);
%
\draw[->,thick,>=stealth] (bench.east) to[out=5,in=180] node[sloped,midway,scale=0.6,above]{Docker API calls} (4,1);
%
\draw[<->,thick,>=stealth] (app1.south) -- node[sloped,midway,scale=0.6,above]{UDP} (app2.north);
\draw[<->,thick,>=stealth] (app1.south) -- node[sloped,midway,scale=0.6,above]{UDP} (appn.north);
\draw[<->,thick,>=stealth] (app2.east) -- node[sloped,midway,scale=0.6,above]{UDP} (appn.west);
%
\draw (8.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\draw (4.1,.75) node[below right,red,scale=0.8] {\textit{Java program}};
\end{tikzpicture}

### 2016-12-27 report/figures/real-churn/bandwidth-figure-real-churn.tex
WIP report

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=1 by 4,
	vertical sep=6mm,
	horizontal sep=0mm,
	xlabels at=edge bottom,
	xticklabels at=edge bottom,
	ylabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
xmax=61,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(0.2,2)},anchor=north, font=\small, draw=none},]
\nextgroupplot[height=2.2cm, 
ylabel={Nodes}, 
ymin=80,ymax=120,
ytick={80,120}, 
yticklabels={80,...,120},
xmajorgrids,
yminorgrids,
tick label style={font=\footnotesize},
label style={font=\tiny}]
\addplot[const plot, color=blue, mark=none] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4}]
\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaverage} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaverage} \closedcycle;
\end{groupplot}
%
\node[anchor=south, rotate=-90] at (group c1r2.east) {EpTO};
\node[anchor=south, rotate=-90] at (group c1r3.east) {JGroups coord};
\node[anchor=south, rotate=-90] at (group c1r4.east) {JGroups nocoord};
\end{tikzpicture}

>>>>

\begin{tikzpicture}
\usetikzlibrary{plotmarks}
\pgfplotsset{
	height=4cm,
	width=\linewidth,
	every axis plot post/.append style={
		solid,
		very thin,
		mark=none
	},
	/pgfplots/area cycle list/.style={/pgfplots/cycle list={%
			{black,fill=black,mark=none},%
			{black,fill=white!25!black,mark=none},%
			{black,fill=white!50!black,mark=none},%
			{black,fill=white!75!black,mark=none},%
			{black,fill=white,mark=none},%
		}
	},
}
\begin{groupplot}[
ymajorgrids,
group style={
	group size=1 by 4,
	vertical sep=6mm,
	horizontal sep=0mm,
	xlabels at=edge bottom,
	xticklabels at=edge bottom,
	ylabels at=edge left,
},
stack plots=y,area style, enlarge x limits=false, 
ymin=0,
xmin=0,
xmax=61,
ylabel={Bandwidth $\left[\SI{}{\mbps}\right]$},
xlabel={Time $\left[\si{\minute}\right]$},
legend columns=5,
legend cell align=left,
legend style={at={(0.6,1.9)},anchor=north, font=\small, draw=none},]
\nextgroupplot[height=2.3cm, 
ylabel={Nodes}, 
ymin=75,ymax=120,
ytick={80,100,120}, 
yticklabels={80,100,120},
xmajorgrids,
yminorgrids,
tick label style={font=\footnotesize},
label style={font=\tiny}]
\addplot[const plot, color=blue, mark=none,set layers,
axis lines = middle,] table [x=minute, y=size, col sep=comma] {csv-data/real-trace.csv};
\nextgroupplot[ymax=4,ytick={0,1,2,3,4},
yticklabels={0,1,2,3,4}]
\addplot table[x=time,y=EpTO-real-trace-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=EpTO-real-trace-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\legend{0, 0.25, 0.5, 0.75, 1}
%
\nextgroupplot%[ymax=4,ytick={0,1,2,3,4},]
\addplot table[x=time,y=JGroups-coord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-coord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
%
\nextgroupplot[ymax=4,ytick={0,1,2,3,4}]
\addplot table[x=time, y=JGroups-nocoord-0.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.250000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.500000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-0.750000, col sep=comma]{\tableaveragereal} \closedcycle;
\addplot table[x=time,y=JGroups-nocoord-1.000000, col sep=comma]{\tableaveragereal} \closedcycle;
\end{groupplot}
%
\node[anchor=south, rotate=-90] at (group c1r2.east) {\epto};
\node[anchor=south, rotate=-90] at (group c1r3.east) {\jgroups-coord};
\node[anchor=south, rotate=-90] at (group c1r4.east) {\jgroups-nocoord};
\end{tikzpicture}

### 2016-12-27 report/figures/real-churn/global-times-real-churn.tex
WIP report

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tableglobaltime};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
\end{axis}
\end{tikzpicture}

>>>>

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
xmin=3649,
xmax=3663,
xtick={3650,3655,3660},
xticklabels={3650,3655,3660},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tableglobaltime};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tableglobaltime};
\legend{\epto, \jgroups-nocoord, \jgroups-coord}
\end{axis}
\end{tikzpicture}

### 2016-12-27 report/figures/real-churn/local-times-real-churn.tex
WIP report

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.1)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tablelocaltime};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tablelocaltime};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaltime};
\legend{EpTO, JGroups-nocoord, JGroups-coord}
\end{axis}
\end{tikzpicture}

>>>>

\begin{tikzpicture}
\begin{axis}[
every axis plot/.append style={very thick},
height=5cm,
grid=major,
grid style={dashed,gray!30},
ymax=1.05,
ymin=0,
xlabel={Time $\left[\si{\second}\right]$},
x tick label style={/pgf/number format/fixed},
ytick={0,0.2,0.4,0.6,0.8,1},
yticklabels={0,0.2,0.4,0.6,0.8,1},
xmin=3649,
xmax=3663,
xtick={3650,3655,3660},
xticklabels={3650,3655,3660},
ylabel=Percentiles,
legend columns=3,
legend cell align=left,
legend style={at={(0.5,1.2)},anchor=north, font=\small, draw=none},
cycle list name=my colors,
]
\addplot+[const plot, mark=none] table[x=EpTO-real-trace-x, y=EpTO-real-trace-y, col sep=comma] {\tablelocaltime};
\addplot+[const plot, mark=none] table[x=JGroups-coord-x, y=JGroups-coord-y, col sep=comma] {\tablelocaltime};
\addplot+[const plot, mark=nonee] table[x=JGroups-nocoord-x, y=JGroups-nocoord-y, col sep=comma] {\tablelocaltime};
\legend{\epto, \jgroups-nocoord, \jgroups-coord}
\end{axis}
\end{tikzpicture}

### 2016-12-27 report/figures/synthetic-churn/bandwidth-figure-churn.tex
Continue work on thesis
WIP report

\begin{tikzpicture}[domain=0:1800]

@@

	width=\linewidth / 2.5,

@@

legend style={at={(1.25,1.5)},anchor=north, font=\small, draw=none},]

@@

\node[anchor=south, rotate=-90] at (group c2r1.east){EpTO};
\node[anchor=south, rotate=-90] at (group c2r2.east){JGroups coord};
\node[anchor=south, rotate=-90] at (group c2r3.east){JGroups nocoord};

>>>>

\begin{tikzpicture}

@@

	width=\linewidth / 2,

@@

legend style={at={(1.15,1.5)},anchor=north, font=\small, draw=none},]

@@

\node[anchor=south, rotate=-90] at (group c2r1.east){\epto};
\node[anchor=south, rotate=-90] at (group c2r2.east){\jgroups-coord};
\node[anchor=south, rotate=-90] at (group c2r3.east){\jgroups-nocoord};

### 2016-12-27 report/framework.tex
WIP report

\section{Framework}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
\subsection{Architecture}
Figure \ref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can broadcast and delivers events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to gather a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the data sent as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. Finally, The framework is open-sourced on Github.
\subsection{Deployment}
\begin{figure}[htp]
	\centering
	\input{figures/architecture.tex}
	\vspace{-2mm} 
	\caption{EpTOTester architecture}
	\vspace{-2mm} 
	\label{fig:complete-architecture}
\end{figure}
To deploy \epto effectively we need two different docker services. One containing the tracker and one containing all \epto replicas. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master mode. This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire EpTOTester architecture is shown in Figure \ref{fig:complete-architecture}. \jt{The figure is heavily inspired from the one in ErasureBench. How should I indicate this?}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}, although the synthetic churn was improved to support adding and removing nodes at the same time.

>>>>

\section{EpToTester}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
\subsection{Architecture}
\autoref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can \epto-broadcast and \epto-deliver events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the data sent as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
\subsection{Deployment}
\begin{figure}[htp]
	\centering
	\input{figures/architecture.tex}
	\vspace{-2mm} 
	\caption{EpTOTester architecture}
	\vspace{-2mm} 
	\label{fig:complete-architecture}
\end{figure}
To deploy \epto effectively we need two different Docker Swarm services. One containing the tracker and one containing all \epto replicas. Both of these services use the same network overlay to communicate. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master mode. This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}. \jt{The figure is inspired from the one in ErasureBench. How should I indicate this?}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}. The synthetic churn provided by it was improved to support adding and removing nodes at the same time.

### 2016-12-27 report/introduction.tex
WIP report

This is where \epto comes into light. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
\epto consists of two distinct parts. The first part is probabilistic dissemination. \epto guarantees that all peers will receive an event with arbitrarily high probability. The second part of \epto is deterministic ordering. Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster id of the events.

@@

The next sections are dedicated to present our benchmarking solution and to present the results obtained. Section \ref{sec:ordering} presents the different kind of ordering. Section \ref{sec:definitions} defines the terms used in the paer. Section \ref{sec:epto} presents \epto and the architecture of our implementation. Section \ref{sec:evaluation} shows and explains the results obtained. We then argue about future work in section \ref{sec:future-work}. We finally draw conclusions in section \ref{sec:conclusion}.

>>>>

This is where \epto shines. \epto, by mixing these two types of algorithms, provides total order along with scalability, validity and integrity. 
\epto consists of two distinct parts. The first part is probabilistic dissemination. \epto guarantees that all peers will receive an event with arbitrarily high probability. The second part of \epto is deterministic ordering. Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events.

@@

The next sections are dedicated to present our benchmarking solution and to present the results obtained. \autoref{sec:ordering} presents the different kinds of ordering. \autoref{sec:definitions} defines the terms used in the paper. \autoref{sec:epto} presents \epto and the architecture of our implementation. \autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced. We then present possible future work in \autoref{sec:future-work}. We finally conclude in \autoref{sec:conclusion}.

### 2016-12-27 report/ordering.tex
WIP report

\section{Ordering Algorithms}
\label{sec:ordering}
\jt{I'm not sure we still need this section}
Distributed systems, like centralized systems, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

>>>>

\section{Ordering Algorithms}
\label{sec:ordering}
\jt{Do we still want this section?}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}
